---
title: "Test1Markdown"
author: "Josh Kostak"
date: "9/26/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes


DATACAMP NOTES

  Intro to R

    class(var) -- checks the class of var
    
    num_vec <- c(1,2,3)
    names(num_vec)<-c("One", "Two", "Three") -- assigning name to elements of vector

    mid <- vector[c(3,4,5)] -- makes a new vector of the the values in vector and positions 3,4, & 5
    
    greater0 <- vector > 0 -- makes a new vector of the values in vector greater then 0
    
    myMatrix <- matrix(1:9, byrow = TRUE, nrow = 3) 
    -- makes a matrix w/3rows that contain numbers 1-9 in row major order

    colnames(matrix)/rownames(matrix) -- name the rows or cols in a matrix
    
    rowSums(matrix) -- sums up rows
    
    plus5 <- vector + 5 
    -- makes a vector of the values in vector plus 5 to each element
    
    factor() -- encode vector as a factor
    
    levels(vector)<- c("Male", "Female") -- creates levels for the vector
    
    summary() -- gives overview of contents
    
    str(dataframe) -- gives you structure of dataframe
    
    dataframe[vector,TRUE] -- selects elements that are true in vector
    
    for(var in vector) -- one way to do for loop
    {
      print(var)
    }
    
    next -- shifts to the next loop iteration
    
    for(i in 1:length(var)) -- another way to do for loop
    {
      print(var[i])
    }

    print(paste("on row", i, "and col", j)) -- concatination
    
    help(____) / ?_____ -- brings up info on fuctions
    
    na.rm -- argument when true removes all empty elements
    
    args(___) -- shows arguments of a function
    
    my_fun <- function(arg1, arg2, ...) -- create your own function
    {
      //body
      return(var)
    }
    
    my_fun <- fuction(a, b=1) -- sets the default of b to be 1
    
    
    
     
  Intro to Data
    
    dataframe$catagoryname <- droplevels(dataframe$catagoryname) 
    -- get rid of whole category in dataframe
    
    email_mutated <- email %>%
        mutate(num_char_cat = ifelse(num_char < med_num_char, "below", "at or below"))
        --creates new col that uses num_char var and simplifies it into 
        "below median" or "at or above median"

    ggplot(data = DF, aes(x=science, y = math, color = subject)) +
      geom_point()      
      --makes a scatterplot and automatically colors based on subject
      
      
      
  Intro to Tidyverse
  
    gapminder %>%
      filter(year == 2007, country = "United States") 
      -- filter gapminder by year = 2007 and country =united states

    gapminder %>%
      arrange(country) -- arrange by country 
      
    arrange(desc(country)) -- same as above but in descending order
      
    gapminder %>%
      mutate(pop = pop/10000) -- modify existing variable
      
    gapminder %>%
      mutate(gdp = gdppercap * pop) -- makes a new variable by multiplying two existing ones together
      
    ggplot(gapminder, aes(x= year, y = lifeExp)) +
      geom_point() +
        scale_x_log10()    --adds log scale to a graph on the x axis
        
    ggplot(gapminder, aes(x= year, y = lifeExp, size = pop)) +
      geom_point()    --makes size of dots represent population variable
      
    facet_wrap(~var) -- add to end of a plot to divide data by var name and display multiple plots
    
    gapminder %>%
      sumarize(meanLifeExp = mean(lifeExp),
                totalpop = sum(pop))      --collapses data down into these vars
                
    gapminder %>%
      group_by(year) %>%
        sumarize(meanLifeExp = mean(lifeExp),
                  totalpop = sum(pop))     --same as above but grouped by year
                  
    group_by(year, pop)   -- you can group by two groups
    
    geom_line(), geom_bar(), geom_histogram(binwidth = 5), geom_boxplot(), geom_density() 
    --other plots, hist only needs x val
    
    geom_bar(position = "dodge") -- makes side by side bar chart
    
    
    
    
   R Markdown
  
    number sign before a word makes it a heading
    spat makes bullet
    word surrounded by two splats makes it bold
    one splt on either side makes it italics
    back ticks means it code
    "[word](url)" -- makes word a link to the url
    dollar signs means its an equation
    back tick r folowed by code and another back tick lets you insert r code
    "```{r chained}" -- name the code chunck chained
    "```{r ref.label = 'chained'}" -- use chained label
     Exploratory Data Analysis
  
    levels(comics$align) -- will show different levelsof align variable
    
    facet_grid(var ~ othervar) -- put first val in rows of grid and 2nd one in cols, add to ggplot
    
    facet_grid(...., labeller = label_both) -- labels vars
    
    
    
    
    
    
CLASS NOTES

  head(_____) -- first 6 elements
  
  tail(_____) -- last 6 elements
  
  ctrl+alt+i -- inserts a code chunk
  
  echo = FALSE -- in markdown -- wont display commands on page
  
  dim(_____) -- dimensions of object
  
  glimpse(___) -- view structure of object
  
  labs(x = "", y = "# of something", title = "Title", subtitle = "subtitle") -- add on end of ggplot
  
  3 s's -- shape, center, spread
  
  facet_grid(.~group) -- another way to use facet, add onto a ggplot
    
   
   
    
    
    
CLASS ASSIGNMENT 

1. In the Flight Delays Case Study in Section 1.1,

    a. The data contain flight delays for two airlines, American Airlines and United Airlines.  Conduct a two-sided permutation test to see if the mean delay times between the two carriers are statistically significant.
    
    Null Hypothesis:
    $H_{0}: \mu_{AA}-\mu_{UA} = 0$
    Verses: 
    $H_{A}: \mu_{AA}-\mu_{UA} \neq 0$
    
    b. The flight delays occured in May and June of 2009.  Conduct a two-sided permutation test to see if the difference in mean delay times between the 2 months is statistically significant.
    
```{r, eval=FALSE}

#Null Hypothesis:
#    $H_{0}: \mu_{AA}-\mu_{UA} = 0$
#    Verses: 
#    $H_{A}: \mu_{AA}-\mu_{UA} \neq 0$

FD <- FlightDelays
glimpse(FD)
 
#find shape of data
ggplot(data=FD, aes(x=Delay)) +
    geom_histogram(color = "black", fill = "purple") +
      labs(title = "BIG TITLE")+
      facet_grid(Carrier~.)

FD %>%
    group_by(Carrier) %>%
      summarize(MeanDelay = mean(Delay), IQRDelay = IQR(Delay), 
                MedianDelay = median(Delay), SDDelay = sd(Delay),
                N = n())

delays <- FD$Delay
#median(delays)
#IQR(delays)
sims <- 10^4 -1

answer <- numeric(sims)

for (i in 1:sims)
{
  #2906 is amount of AA delays and 4029 is the total number of delays
  index <- sample(4029, 2906, replace = FALSE)
  answer[i] <- mean(delays[index]) - mean(delays[-index])
}

obs <-tapply(FD$Delay, FD$Carrier, mean)
obs
obs_diff <- obs[1] - obs[2] 
#obs[1] is the first carriers mean delay (AA) and obs[2] is the second one (UA)
obs_diff

pval <- (sum(answer <= obs_diff)+1)/(sims+1)
pval
``` 

> SOLUTION: 

```{r, eval=FALSE}
FD <- FlightDelays
glimpse(FD)
FD %>%
    group_by(Month) %>%
      summarize(m = n(), MeanDelay = mean(Delay))

delays <- FD$Delay
sims <- 10^4 -1

answer <- numeric(sims)

#mixes data up
for (i in 1:sims)
{
  index <- sample(4029, 2030, replace = FALSE)
  answer[i] <- mean(delays[index]) - mean(delays[-index])
}

#applies mean of both months
obs <-tapply(FD$Delay, FD$Month, mean)
obs
#finds difference in means
obs_diff <- obs[1] - obs[2]
obs_diff

#if the val in answer is less then the diff then add one and 
#divide by number of simulation element
#idk where the < comes from...
pval <- (sum(answer < obs_diff)+1)/(sims+1)
pval
```

> SOLUTION: 
    
2. In the Flight Delays Case Study in Section 1.1, the data contain flight delays for two airlines, American Airlines and United Airlines.

    a. Compute the proportion of times that each carrier's flights was delayed more than 20 minutes.  Conduct a two-sided test to see if the difference in these proportions is statistically significant.
    
    b. Compute the variance in the flight delay lengths for each carrier.  Conduct a test to see if the variance for United Airlines is greater than that of American Airlines.
    
        the null hypo is (signma of UA squared)/(sigma of AA) = 1
    
```{r, eval=FALSE}
# a. Your code here
FD <- FlightDelays
glimpse(FD)
FD %>%
    group_by(Carrier) %>%
      summarize(m = n(), MeanDelay = mean(Delay > 20))

delays <- FD$Delay
sims <- 10^4 -1

answer <- numeric(sims)

#mixes data up
for (i in 1:sims)
{
  index <- sample(4029, 2906, replace = FALSE)
  answer[i] <- mean(delays[index] > 20) - mean(delays[-index] > 20)
}

#applies mean of both months
obs <-tapply(FD$Delay > 20, FD$Carrier, mean)
obs
#finds difference in means
obs_diff <- obs[1] - obs[2]
obs_diff

#if the val in answer is less then the diff then add
#one and divide by number of simulation element
#idk where the < comes from...
pval <- (sum(answer < obs_diff)+1)/(sims+1)
pval
```

> SOLUTION: 

```{r, eval=FALSE}
#this is the one you must divide the stuff instead of subtract
# b. Your code here
FD <- FlightDelays
glimpse(FD)
FD %>%
    group_by(Carrier) %>%
      summarize(m = n(), VarienceDelay = var(Delay))

delays <- FD$Delay
sims <- 10^4 -1

answer <- numeric(sims)

#mixes data up
for (i in 1:sims)
{
  index <- sample(4029, 2906, replace = FALSE)
  answer[i] <- var(delays[index]) / var(delays[-index])
}

#applies mean of both months
obs <-tapply(FD$Delay, FD$Carrier, var)
obs
#finds difference in means
obs_diff <- obs[1] / obs[2]
obs_diff

#if the val in answer is less then the diff then add one and 
#divide by number of simulation element
#idk where the < comes from...
pval <- (sum(answer < obs_diff)+1)/(sims+1)
pval
```




Class assignment from tuesday:
```{r, eval = FALSE}
library(readxl)
library(dplyr)
library(ggplot2)


DF <- read_excel("TMP.xlsx")

#fix errors in the data set created by excel
DF <- DF %>%
      mutate(Age_Cohort = gsub("12-Jun", "6-12", Age_Cohort))

DF <- DF %>%
  mutate(Age_Cohort = gsub("42898", "6-12", Age_Cohort))

DF <- DF %>%
  mutate(Age_Cohort = gsub("0 - 5", "0-5", Age_Cohort))

DF


#start finding answers for the questions on the assignment
DF %>%
  filter(Gender == "Male") %>%
  summarize(MeanMaleExpenditures = mean(Expenditures))

male <- 18001

DF %>%
  filter(Ethnicity == "Hispanic") %>%
    summarize(MeanHispanicExpenditures = mean(Expenditures))

hispanic <- 11066

DF %>%
  filter(Age_Cohort == "22-50") %>%
  summarize(Mean22to50Expenditures = mean(Expenditures))

twentytwotofifty <- 40209

DF %>%
  filter(Age_Cohort == "22-50", Ethnicity == "White not Hispanic", Gender == "Male") %>%
  summarize(MeanMW22to50Expenditures = mean(Expenditures))

whiteMale22to50 <- 38604

DF %>%
  filter(Age_Cohort == "22-50", Ethnicity == "Asian") %>%
  summarize(MeanAsian22to50Expenditures = mean(Expenditures))

asian22tofifty <- 39581


#make a dataframe from the results and turn dataframe into a bar chart
bars <- data.frame(Catagory = c("Male","Hispanic","22-50", "White Male 22-50", "Asian 22-50"), 
                  values = c(male, hispanic, twentytwotofifty, whiteMale22to50, asian22tofifty))

ggplot(bars, aes(x=Catagory, weight=values)) + 
  geom_bar() +
  labs(x = "Catagory", y = "Mean Expenditures", title = "Average Expenditures")




#instructor example of grouping by gender and piping into ggplot
DF %>% 
  group_by(Gender) %>% 
  summarize(ME = mean(Expenditures), MDE = median(Expenditures), n= n()) %>%
  ggplot(aes(x = Gender, y= ME, fill = Gender)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Average Expenditure by Gender", y = "Mean Expenditure") + 
  theme_bw() + 
  scale_fill_manual(values = c("pink", "blue"))

#instructor example of grouping by ethnicity and piping into ggplot
DF %>% 
  group_by(Ethnicity) %>% 
  summarize(ME = mean(Expenditures), MDE = median(Expenditures), n= n()) %>%  
  ggplot(aes(x = reorder(Ethnicity, ME), y = ME)) +
  geom_bar(stat="identity", fill = "red") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) + 
  labs(x = "", y = "Mean Expenditure", title = "Average Expenditure by Ethnicity")


```